A molecular taxonomy of tumors independent of tissue-of-origin

### 2021-07-02 text/introduction.Rmd
Minor edits to text

More recently, pan-cancer mutations and other data have been looked at as
disruptions of normal pathway activity. Pathway-centric analysis of tumors
provides additional benefits over gene-centric analysis: (1) It is significantly 
less noisy because it aggregates molecular events across multiple genes in the 
same pathway and (2) the identification of a potential casual mechanism is easier 
to interpret as genetic aberrations are linked to molecular pathways. In a perspective, 
Creixell et al. [@Creixell-pathwaynetworkanalysis] described approaches of pathway 
and network analysis applicable to next-generation sequencing data. The most common 
approach being fixed-gene enrichment analysis to identify over-represented pathways. 
Sanchez-Vega et al. [@PCAWG-PATH] carried out an exhaustive
analysis of pathway enrichment using whole genome data. Focusing on ten frequently 
altered canonical signaling pathways in cancer, they explored the mutation frequencies 
of coding regions of genes in these pathways for each cancer type. Network approaches
address the sparseness of mutations, allowing genes to be influenced by
mutations in nearest network neighbors [@hotnet2;@hornNetSigNetworkbasedDiscovery2018;
@Iranzo-Network]. These methods are powerful especially when applied to smaller
numbers of tumors. One of the challenges in the field remains that there may be
thousands of moderate effect genes that occur at such low frequency that they
are impossible to detect using positive selection theory. Some researchers have
attempted to address this challenge by applying machine learning to cancer data
to discover groups of functionally related genes as they interact with larger
pathways and networks [@kimPancancerAnalysisSomatic2018;
@mourikis-ml-pathways; @moonlight]. These studies have proven very effective at
highlighting fundamental disease phenotypes at the pathway level across cancers
with different origins at the cellular and tissue level.

>>>>

More recently, pan-cancer mutations and other data have been looked at as
disruptions of normal pathway activity. Pathway-centric analysis of tumors
provides additional benefits over gene-centric analysis: (1) It is significantly 
less noisy because it aggregates molecular events across multiple genes in the 
same pathway and (2) the identification of a potential casual mechanism is easier 
to interpret as genetic aberrations are linked to molecular pathways. In a perspective, 
Creixell et al. [@Creixell-pathwaynetworkanalysis] described approaches of pathway 
and network analysis applicable to next-generation sequencing data, the most common 
approach being fixed-gene enrichment analysis to identify over-represented pathways. 
Sanchez-Vega et al. [@PCAWG-PATH] carried out an exhaustive
analysis of pathway enrichment using whole genome data. Focusing on ten frequently 
altered canonical signaling pathways in cancer, they explored the mutation frequencies 
of coding regions of genes in these pathways for each cancer type. Network approaches
address the sparseness of mutations, allowing genes to be influenced by
mutations in nearest network neighbors [@hotnet2;@hornNetSigNetworkbasedDiscovery2018;
@Iranzo-Network]. These methods are powerful especially when applied to smaller
numbers of tumors. One of the challenges in the field remains that there may be
thousands of moderate effect genes that occur at such low frequency that they
are impossible to detect using positive selection theory. Some researchers have
attempted to address this challenge by applying machine learning to cancer data
to discover groups of functionally related genes as they interact with larger
pathways and networks [@kimPancancerAnalysisSomatic2018;
@mourikis-ml-pathways; @moonlight]. These studies have proven very effective at
highlighting fundamental disease phenotypes at the pathway level across cancers
with different origins at the cellular and tissue level.

### 2021-07-02 text/survival.Rmd
Minor edits to text

We found that cancer types, as expected, have a range of prognoses relative to
the general population. In **Fig 5B-D** we show how effective age baseline, and
effective age rate (which modifies the baseline based on your actual age) as
well as the $k_{tis}$ which accelerates or decelerates further aging, varies
considerably from cancer to cancer. For example a 20 year old diagnosed with
THCA or PRAD would be treated similarly to a 40 year old based on the Effective
Age parameter. It's noticeable that for every year beyond 20 in the PRAD
diagnosis we add only an extra ~ 0.2 years (Effective Age Rate). The $k_{tis}$
parameter represents an acceleration (greater than 1) or deceleration (less) of
the further aging beyond diagnosis. For example for BRCA patients each year past
diagnosis adds about 1 year of risk, whereas for LGG patients it adds around 2,
and for SARC patients it adds only 0.25 or so.

>>>>

We found that cancer types, as expected, have a range of prognoses relative to
the general population. In **Fig 5B-D** we show how effective age baseline, and
effective age rate (which modifies the baseline based on your actual age) as
well as the $k_{tis}$ which accelerates or decelerates further aging, varies
considerably from cancer to cancer. For example a 20 year old diagnosed with
THCA or PRAD would be treated similarly to a 40 year old based on the Effective
Age parameter. It's noticeable that for every year beyond 20 in the PRAD
diagnosis we add only an extra ~ 0.2 years (Effective Age Rate). The $k_{tis}$
parameter represents an acceleration (greater than 1) or deceleration (less) of
the further aging beyond diagnosis. For example, for BRCA patients each year past
diagnosis adds about 1 year of risk, whereas for LGG patients it adds around 2,
and for SARC patients it adds only 0.25 or so.
